---
title: "UFOlogy - UFOs' observations in UK from 1974 to 2014"
author: "Sylvain Deniau"
date: "30/10/2022"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Presentation & data used

The main objective of this application is to show the location and type of different ufological phenomena that have been observed in the United Kingdom from 1974 to 2014.

The original dataset (https://www.kaggle.com/datasets/camnugent/ufo-sightings-around-the-world) does contain over 80,000 worldwide records of UFO sightings dating back as far as 1949 with  latitude and longitude data, dates, duration and description of the phenomenom.

`global.R` download the global dataset and provide some treatment in order to filter the data in particuliar to select location (i.e. UK) and extract year from Date_time variable.

```{r eval = FALSE, echo=TRUE}
ufo <- ufodata %>% select(Date_time, UFO_shape, country, latitude, longitude, city) %>%
        filter(country=="gb") %>% 
        filter(!is.na(Date_time)&!is.na(UFO_shape)&!is.na(latitude)&!is.na(longitude)) %>% 
        filter(UFO_shape!=""&latitude!=""&longitude!="") %>% 
        mutate(time=year(mdy_hm(Date_time)),latitude=as.numeric(latitude)) %>%
        mutate(city=str_to_title(city)) %>% 
        filter(!is.na(latitude)) %>% select(-Date_time)
```



## Location of the various files

- `global.R`, `ui.R` and `server.R` can be found there https://github.com/sdeniau/devdataproducts

- The **Shiny App** is to be found here https://sdeniau.shinyapps.io/UFOlogy/

- The *Rmd* file used to do this presentation is to be find there https://github.com/sdeniau/devdataproducts/prezufology.Rmd

## Calculation

Calculation done in `server.R` are simple.

On the one side it filter the data set with the selected year and create a Leaflet output 
```{r ,echo = TRUE, eval=FALSE}

output$ufomap <- renderLeaflet({
        ufoselect <- ufo %>% filter(time==input$year)
        ufoselect %>% 
            leaflet() %>%
            addTiles() %>%
            addMarkers(lat=~latitude, lng=~longitude, label=~city)

    })

```

On the other side it filter the data set with the selected year and calculate the number of UFOs' observation per shape, sort them in a descending order and create a Table output

```{r ,echo = TRUE, eval=FALSE}

 output$ufotable <- renderTable({
            ufotable <- ufo %>% filter(time==input$year) %>% 
                    group_by(UFO_shape) %>% summarise (nb=n()) %>% 
                    arrange(desc(nb))
            ufotable
    })

```

## Illustration 

For example for the year 1995 the map will look like:


```{r message = FALSE, warning = FALSE}
        
library(tidyverse)
library(lubridate)
library(leaflet)
library(shiny)
library(shinyhelper)


id <- "1cTekOyvQa0sRKp9IChFW232N67hFidAm"
ufodata <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", id))


ufo <- ufodata %>% select(Date_time, UFO_shape, country, latitude, longitude, city) %>%
        filter(country=="gb") %>% 
        filter(!is.na(Date_time)&!is.na(UFO_shape)&!is.na(latitude)&!is.na(longitude)) %>% 
        filter(UFO_shape!=""&latitude!=""&longitude!="") %>% 
        mutate(time=year(mdy_hm(Date_time)),latitude=as.numeric(latitude)) %>%
        mutate(city=str_to_title(city)) %>% 
        filter(!is.na(latitude)) %>% select(-Date_time) 

ufoselect <- ufo %>% filter(time==1995)
        ufoselect %>% 
            leaflet() %>%
            addTiles() %>%
            addMarkers(lat=~latitude, lng=~longitude, label=~city)
```

